# Getting Started

## Introduction

This chapter provides a quick tutorial on how to install and set up R and Stata on both Windows and Mac computers. By the end of this chapter, you'll have the necessary tools ready to begin your analysis.

## Installing R

### Windows

1.  **Download R**:
    -   Go to the [R Project website](https://cran.r-project.org/).
    -   Click on "Download R for Windows."
    -   Click on "base" to download the base R package.
2.  **Install R**:
    -   Run the downloaded `.exe` file.
    -   Follow the installation instructions, accepting the default settings.
3.  **Install RStudio** (Optional but recommended):
    -   Download RStudio from the [RStudio website](https://rstudio.com/products/rstudio/download/).
    -   Run the installer and follow the setup instructions.

### Mac

1.  **Download R**:
    -   Visit the [R Project website](https://cran.r-project.org/).
    -   Click on "Download R for macOS."
2.  **Install R**:
    -   Open the downloaded `.pkg` file.
    -   Follow the installation instructions.
3.  **Install RStudio** (Optional but recommended):
    -   Download RStudio from the [RStudio website](https://rstudio.com/products/rstudio/download/).
    -   Open the `.dmg` file and drag RStudio to your Applications folder.

## Installing Stata

### Windows

1.  **Obtain a License**:
    -   Stata is (unfortunately) commercial software. You need a valid license. Look it up in your course page or ask your teacher.
2.  **Download Stata**:
    -   Go to the [Stata website](https://www.stata.com/) and log in to your account to download the installer.
3.  **Install Stata**:
    -   Run the downloaded `.exe` file.
    -   Follow the installation instructions, entering your license information when asked for it.

### Mac

1.  **Obtain a License**:
    -   Stata is (unfortunately) commercial software. You need a valid license. Look it up in your course page or ask your teacher.
2.  **Download Stata**:
    -   Visit the [Stata website](https://www.stata.com/) and log in to your account to download the installer.
3.  **Install Stata**:
    -   Open the downloaded `.dmg` file.
    -   Drag the Stata application to your Applications folder.
    -   Launch Stata and enter your license information.

## Setting Up Your Environment

### R Setup

1.  **Open RStudio** (or R GUI if not using RStudio).
2.  **Install Essential Packages**:
    -   Open the Console and run:

```{r eval=FALSE}
   install.packages(c("tidyverse", "magrittr", "here"))
```

3.  **Create a New Project** (Optional but recommended in RStudio):
    -   Go to "File" \> "New Project" \> "New Directory" \> "New Project."
    -   Choose a location and name for your project, then click "Create Project."

### Stata Setup

1.  **Open Stata**.

2.  **Set a Working Directory**:

    -   Use the command:

    ```{stata eval=FALSE}
       cd "path/to/your/directory"
    ```

    Replace `"path/to/your/directory"` with the path where you want to save your files.

3.  **Creating Do-Files**:

    -   Go to "File" \> "New Do-file Editor."
    -   Save the Do-file in your working directory.

## Verification

### R

1.  **Test Installation**:

    -   In RStudio or R GUI, type:

    ```{r eval=FALSE}
       print("R is working! Yay!")
    ```

    -   If you see the output `[1] "R is working!"`, your installation is successful.

2.  **Load a Package**:

    -   Run:

    ```{r eval=FALSE}
       library(ggplot2)
       print("Package 'ggplot2' is loaded!")
    ```

### Stata

1.  **Test Installation**:

    -   In the Command window, type:

    ```{stata eval=FALSE}
       display "Stata is working! Yay!"
    ```

    -   If you see the output `Stata is working!`, your installation is successful.

2.  **Check Version**:

    -   Type:

    ```{stata eval=FALSE}
       about
    ```

    -   This will display the version of Stata installed.

------------------------------------------------------------------------

With your environment set up, you're now ready to start performing analyses using R and Stata!
