# Linear Regression

## Introduction
This chapter covers how to perform linear regression to study the relationship between variables. We'll use an example dataset that simulates the relationship between study time and performance on an online learning platform.

## Example Question
**How does the amount of time spent on an e-learning platform (in hours) affect the test scores of users?**

## Dataset Simulation in R

```{r eval=FALSE}
# Load necessary package
set.seed(123)

# Simulate data
n <- 100
study_time <- rnorm(n, mean = 10, sd = 2)  # Average 10 hours
test_score <- 50 + 5 * study_time + rnorm(n, mean = 0, sd = 5)  # Linear relationship with some noise

# Create a data frame
data <- data.frame(study_time, test_score)

# View the first few rows
head(data)
```

## Dataset Simulation in Stata

```{stata eval=FALSE}
* Set seed for reproducibility
set seed 123

* Simulate data
set obs 100
gen study_time = rnormal(10, 2)
gen test_score = 50 + 5 * study_time + rnormal(0, 5)

* View the first few rows
list in 1/10
```

## Performing Linear Regression

### R

```{r eval=FALSE}
# Fit the linear regression model
model <- lm(test_score ~ study_time, data = data)

# View the summary
summary(model)
```

### Stata

```{stata eval=FALSE}
* Fit the linear regression model
regress test_score study_time
```

## Assumptions
- **Linearity**: The relationship between the independent and dependent variable should be linear.
- **Independence**: Observations should be independent of each other.
- **Homoscedasticity**: The residuals should have constant variance at every level of the independent variable.
- **Normality**: The residuals should be normally distributed.
