# Logistic Regression

## Introduction
This chapter covers logistic regression, which is used when the outcome variable is binary. The example dataset will examine whether the frequency of technical support contact predicts whether a user continues to use a software product.

## Example Question
**Does the frequency of contacting technical support predict whether a user will continue using a software product?**

## Dataset Simulation in R

```{r eval=FALSE}
# Load necessary package
set.seed(123)

# Simulate data
n <- 200
support_contact <- rpois(n, lambda = 2)  # Number of contacts with support
continued_use <- rbinom(n, size = 1, prob = 1 / (1 + exp(-(-1 + 0.5 * support_contact))))

# Create a data frame
data <- data.frame(support_contact, continued_use)

# View the first few rows
head(data)
```

## Dataset Simulation in Stata

```{stata eval=FALSE}
* Set seed for reproducibility
set seed 123

* Simulate data
set obs 200
gen support_contact = rpoisson(2)
gen continued_use = rbinomial(1, 1 / (1 + exp(-(-1 + 0.5 * support_contact))))

* View the first few rows
list in 1/10
```

## Performing Logistic Regression

### R

```{r eval=FALSE}
# Fit the logistic regression model
model <- glm(continued_use ~ support_contact, data = data, family = "binomial")

# View the summary
summary(model)
```

### Stata

```{stata eval=FALSE}
* Fit the logistic regression model
logit continued_use support_contact
```

## Assumptions
- **Binary Outcome**: The dependent variable should be binary.
- **Independence**: Observations should be independent of each other.
- **Linearity of logit**: The logit (log-odds) of the outcome should be linearly related to the predictors.
- **No multicollinearity**: The predictors should not be highly correlated with each other.
- **Large sample size**: Logistic regression typically requires a large sample size to provide reliable estimates.
