# Multilevel Regression

## Introduction

This chapter covers multilevel regression, where data is nested. We will explore how user satisfaction with a mobile app is affected by time spent on the app, considering that users are nested within different age groups.

## Example Question

**Does time spent on a mobile app influence user satisfaction, and does this effect differ across age groups?**

## Dataset Simulation in R

```{r eval=FALSE}
# Load necessary package
set.seed(123)

# Simulate data
n_groups <- 5  # Number of age groups
n_per_group <- 50  # Number of users per group

age_group <- factor(rep(1:n_groups, each = n_per_group))
time_spent <- rnorm(n_groups * n_per_group, mean = 30, sd = 10)
satisfaction <- 3 + 0.2 * time_spent + as.numeric(age_group) + rnorm(n_groups * n_per_group, sd = 2)

# Create a data frame
data <- data.frame(age_group, time_spent, satisfaction)

# View the first few rows
head(data)
```

## Dataset Simulation in Stata

```{stata eval=FALSE}
* Set seed for reproducibility
set seed 123

* Simulate data
set obs 250
gen group = ceil(_n/50)  // Age group
gen time_spent = rnormal(30, 10)
gen satisfaction = 3 + 0.2 * time_spent + group + rnormal(0, 2)

* Convert group to a factor
egen group_factor = group(group)

* View the first few rows
list in 1/10
```

## Performing Multilevel Regression

### R

```{r eval=FALSE}
# Load necessary package
library(lme4)

# Fit the multilevel model
model <- lmer(satisfaction ~ time_spent + (1 | age_group), data = data)

# View the summary
summary(model)
```

### Stata

```{stata eval=FALSE}
* Fit the multilevel model
mixed satisfaction time_spent || group:
```

## Assumptions

-   **Normality of residuals**: The residuals at each level of the model should be normally distributed.
-   **Linearity**: The relationship between predictors and the outcome should be linear at each level of the model.
-   **Independence**: Observations within each group should be independent.
-   **Homoscedasticity**: The variance of residuals should be consistent across all levels of the hierarchy.
